---
title: "GA_Elections"
output: html_document
date: "2025-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Libraries 

```{r}
library(tidyverse)
library(tidyr)
library(readxl)
library(stringr)
library(pdftools)
```


# Define functions
```{r}
cleanpage <- function(page) {
  ## This function takes a list of data from a page of a pdf and indexes 
  ## the list to the desired data 
  # Inputs: 
    ## page = a page from a pdf parsed as a list of vectors 
  # Output 
    ## The data indexed to start at the line below the variable names and ending before text
  
  n <- 0
  m <- 0
  
  for (i in 1:length(page)) {
    
    
    val <- grepl("County", page[i])
     
    if(val) { 
      n <- i+1
      #print(n)
      
    }
    
    
    val2 <- grepl("Report prepared", page[i])
    if(val2) { 
      m <- i-1
      #print(m)
      
    }
  }
  
  #print(paste(n," ",m))
  return(page[n:m])
}
```



# Load data 

## - Elections data 

```{r}
#getwd()
path <- "./GA_2022Nov.csv"
data <- read.csv(path)

```


## - County data 

"County Data" Source: https://www.legis.ga.gov/joint-office/reapportionment 
- Downloaded pdfs containing *2021* data on the state senate, state house, and congressional districts contained by each county in Georgia.

According to this *Article*, the 2021 maps were used in the 2022 elections despite being contested in court by pro-democracy groups.
https://www.democracydocket.com/analysis/nearly-two-years-later-georgias-congressional-map-heads-to-trial/ 


Court District sources 
- https://gapubdef.org/store-category/offices-by-judicial-circuit/  
- https://www.gasd.uscourts.gov/court-info
- https://www.southernjudicialcircuit.com/Georgia-Circuits.pdf 

```{r}

path2 <- "./GA_housedistricts_2021.pdf"
house <- pdf_text(path2)

path3 <- "./GA_senatedistricts_2021.pdf"
senate <- pdf_text(path3)
  
path4 <- "./GA_congressdistricts_2021.pdf"
congress <- pdf_text(path4)

```


# Clean pdf data

## - House Districts 
```{r}
# Split into pages
test <- str_split(house, "\n")

# Apply clean page function 
for (i in 1:length(test)) {test[[i]] <- cleanpage(test[[i]])}


## Test 
result <- cleanpage(test[[1]])

```


```{r}

county_list <- list()
district_list <- list()

for (i in 1:length(test)) {
  
  for (j in 1:length(test[[i]])) {
    
    cnty <- str_remove_all(str_extract_all(test[[i]][j], "^\\s*(?:\\S+\\s+)"), " ") # Extract first word from string 
    dstr <- str_extract_all(test[[i]][j], "([0-9]+)")[[1]][1] # Extract first number from string
    
    county_list <- append(county_list, cnty)
    district_list <- append(district_list, dstr)
  }
}


county_list <- unlist(county_list)
district_list <- unlist(district_list)
```

```{r}
county_district_key <- data.frame(County_name = county_list, District_name = district_list)
county_district_key <- drop_na(county_district_key)
```



## - Senate Districts

```{r}
# Split into pages
sen <- str_split(senate, "\n")

# Apply clean page function 
for (i in 1:length(sen)) {sen[[i]] <- cleanpage(sen[[i]])}


```

```{r}
county_list2 <- list()
district_list2 <- list()

for (i in 1:length(sen)) {
  
  for (j in 1:length(sen[[i]])) {
    
    cnty <- str_remove_all(str_extract_all(sen[[i]][j], "^\\s*(?:\\S+\\s+)"), " ") # Extract first word from string 
    dstr <- str_extract_all(sen[[i]][j], "([0-9]+)")[[1]][1] # Extract first number from string
    
    county_list2 <- append(county_list2, cnty)
    district_list2 <- append(district_list2, dstr)
  }
}


county_list2 <- unlist(county_list2)
district_list2 <- unlist(district_list2)

```

```{r}
county_district_key2 <- data.frame(County_name = county_list2, District_name = district_list2)
county_district_key2 <- drop_na(county_district_key2)
```

## - Congressional Districts

```{r}
# Split into pages 
con <- str_split(congress, "\n")

# Apply clean page function 
for (i in 1:length(con)) {con[[i]] <- cleanpage(con[[i]])}
```

```{r}
county_list3 <- list()
district_list3 <- list()

for (i in 1:length(con)) {
  
  for (j in 1:length(con[[i]])) {
    
    cnty <- str_remove_all(str_extract_all(con[[i]][j], "^\\s*(?:\\S+\\s+)"), " ") # Extract first word from string 
    dstr <- str_extract_all(con[[i]][j], "([0-9]+)")[[1]][1] # Extract first number from string
    
    county_list3 <- append(county_list3, cnty)
    district_list3 <- append(district_list3, dstr)
  }
}

county_list3 <- unlist(county_list3)
district_list3 <- unlist(district_list3)

```

```{r}
county_district_key3 <- data.frame(County_name = county_list3, District_name = district_list3)
county_district_key3 <- drop_na(county_district_key3)
```


# Clean elections data 

```{r}
# Separate names from parties 
party_list <- list()

for (i in 1:nrow(data["choice.name"])) {
  
  name <- data["choice.name"][i,]
  data["choice.name"][i,] <- unlist(str_split(name, "[(]"))[1]
  party_list[i] <- str_extract(name, "[ ]\\(.*")
  
  #unlist(str_split(name, "[(]"))[2]
}


# Make candidate party 
data["Candidate_party"] <- unlist(party_list)

# Make candidate party standard 
data <- data %>% 
  mutate(Candidate_party = case_when(grepl("Rep", Candidate_party)==T ~"REP", 
                                            grepl("Dem", Candidate_party)==T ~"DEM",
                                            grepl("Lib", Candidate_party)==T~"LIB",
                                            grepl("Dem", Candidate_party)==F & grepl("Rep",
                                                                                    Candidate_party)==F & grepl("Lib", Candidate_party)==F~"OTE"))
```



```{r}
# Rename columns 
data <- data %>% rename(Contest_name = contest.name, 
                        Candidate_name = choice.name)
```


```{r}
# Make Contest_name standard - remove the "Vote for 1" section
data["Contest_name"] <- 
  unlist(lapply(data["Contest_name"], function(x) str_remove(x, "[ ]\\(.*"))[[1]]) 
```


```{r}
# Select choice columns
data <- data %>% 
  select(Contest_name, Candidate_name, Candidate_party)

```


# Make County Keys

## - State house races 
```{r}
# Required: county_district_key, data 

## Subset election data to state house races
  state_house_races <- data %>% select(Contest_name) %>% unique() %>% 
    filter(grepl("State House", Contest_name)==T) %>% pull()


## Make a contest-district key 
  ckey1 <- data.frame(Contest_name = state_house_races, 
                   District_name = as.vector(sapply(state_house_races, function(x)
                     str_extract_all(x, "([0-9]+)")[[1]][1])))

## Join the contest-district key to the county-district key
  state_house_key <- left_join(ckey1, county_district_key, by = "District_name")

```





## - State Senate races 
```{r}
# Required: county_district_key2, data

## Subset election data to state house races
  state_senate_races <- data %>% select(Contest_name) %>% unique() %>% 
    filter(grepl("State Senate", Contest_name)==T) %>% pull()

## Make a contest-district key 
  ckey2 <- data.frame(Contest_name = state_senate_races, 
                   District_name = as.vector(sapply(state_senate_races, function(x)
                     str_extract_all(x, "([0-9]+)")[[1]][1])))

## Join the contest-district key to the county-district key
  state_senate_key <- left_join(ckey2, county_district_key2, by = "District_name")

```


## - Congressional races 
```{r}
# Required: county_district_key3, data

## Subset election data to state house races
  state_congress_races <- data %>% select(Contest_name) %>% unique() %>% 
    filter(grepl("US House", Contest_name)==T) %>% pull()

## Make a contest-district key 
  ckey3 <- data.frame(Contest_name = state_congress_races, 
                   District_name = as.vector(sapply(state_senate_races, function(x)
                     str_extract_all(x, "([0-9]+)")[[1]][1])))


## Join the contest-district key to the county-district key
  state_congress_key <- left_join(ckey3, county_district_key3, by = "District_name")

```



## - District Attorney cases 

```{r}
state_district_races <- data %>% select(Contest_name) %>% 
  filter(grepl("District Attorney", Contest_name)==T) %>% unique() %>% pull()
```


```{r}
# SOURCE TO MANUALLY FILL IN COUNTIES 
## https://gapubdef.org/store-category/offices-by-judicial-circuit/ 

# Waycross - 1st 
  Waycross <- toupper(str_split("Bacon, Brantley, Charlton, Coffee, Pierce, Ware", ", ")[[1]])

# Atlantic - 1st 
  Atlantic <- toupper(str_split("Bryan, Evans, Liberty, Long, McIntosh, Tattnall", ", ")[[1]])

# Alapaha - 2nd 
  Alapaha <- toupper(str_split("Atkinson, Berrien, Clinch, Cook, Lanier", ", ")[[1]])

# Pataula - 2nd 
  Pataula <- toupper(str_split("Clay, Early, Miller, Quitman, Randolph, Seminole, Terrell", ", ")[[1]])

# Cherokee - 7th 
  Cherokee <- toupper(str_split("Bartow, Gordon", ", ")[[1]])

# Columbia - 
  Columbia <- toupper(str_split("Columbia", ", ")[[1]])

# Cordele - 8th
  Cordele <- toupper(str_split("Ben Hill, Crisp, Dooly, Wilcox", ", ")[[1]])

# Enotah - 9th 
  Enotah <- toupper(str_split("Lumpkin, Towns, Union, White", ", ")[[1]])

# Northeastern - 9th
  Northeastern <- toupper(str_split("Dawson, Hall", ", ")[[1]])

# Paulding - 7th
  Paulding <- toupper(str_split("Paulding", ", ")[[1]])

# Rockdale - 4th
  Rockdale <- toupper(str_split("Rockdale", ", ")[[1]])

```


```{r}
# Make a Contest-County key 

## Required: state_district_races

circuit_list <- list()
county_list <- list()

for (i in 1:length(state_district_races)) {
  
  circuit <- state_district_races[i]
  circuit <- str_remove(circuit, "District Attorney")
  circuit <- str_remove(circuit, "Judicial Circuit")
  circuit <- str_remove(circuit, "-")
  circuit <- str_trim(circuit)
  
  clist <- eval(parse(text = circuit)) # Evaluate the variables generated in block above -- names match up with circuits
  num <- length(clist)
  
  circuit_list <- append(circuit_list, rep(state_district_races[i], num)) # Repeat Contest_name for each county 
  county_list <- append(county_list, clist) # Append counties 
  
}


# County - Office key for Congressional offices 
state_district_key <- data.frame(County_name = unlist(county_list),
                               Contest_name = unlist(circuit_list))

```


## - Statewide races 

```{r}
# Get list of all counties 
GA_counties <- state_congress_key %>% select(County_name) %>% unique() %>% pull()

# Filter for statewide races 
statewideraces <- data %>% select(Contest_name) %>% unique() %>% 
    filter(!(Contest_name %in% state_congress_races) & 
           !(Contest_name %in% state_senate_races) & 
           !(Contest_name %in% state_house_races)) %>% 
    filter(grepl("District", Contest_name)==F & 
           grepl("Amendment", Contest_name)==F & 
           grepl("Referendum", Contest_name)==F) %>% pull()

```


```{r}
office_list <- list()

# Repeat counties for each statewide race 
  for (i in 1:length(statewideraces)) {
    office_list <- append(office_list, rep(statewideraces[i],length(GA_counties)))
  }

# County - Office key for Congressional offices 
  statewide_office_key <- 
    data.frame(Contest_name = unlist(office_list),
               County_name = rep(GA_counties,
                                 length(statewideraces)))
```



## RBind keys 

```{r}
# Required: statewide_office_key, state_congress_key, state_senate_key, state_house_key

## Get rid of district number if needed
state_congress_key <- state_congress_key %>%
  select(-District_name)
state_senate_key <- state_senate_key %>% 
  select(-District_name)
state_house_key <- state_house_key %>% 
  select(-District_name)

## RBind all the keys
county_key <- rbind(state_congress_key, state_senate_key, state_house_key, statewide_office_key, state_district_key)

```









