---
title: "WI_Data"
output: html_document
date: "2025-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries 

```{r}
library(tidyverse)
library(tidyr)
library(readxl)
library(stringr)
```


# TO DO 
#---- Add 50 rows for each statewide race 
#---- Make a statewide County-race key? 
#---- the statewide data has state rep and senator races...need to avoid duplicates


# Define functions 

```{r}
# Code from 
# https://www.geeksforgeeks.org/how-to-read-a-xlsx-file-with-multiple-sheets-in-r/
read_all_sheets <- function(file) {
  
  # get info about sheets 
  sheets <- excel_sheets(file)
  sheets <- sheets[-1] # Get rid of "Document map"
  tibble <- lapply(sheets, function(x) read_excel(file, sheet = x))
  data_frame <- lapply(tibble, as.data.frame)
  
  # assigning names to data frames 
  names(data_frame) <- sheets
  
  # print data frame 
  return(data_frame)
}

```

# Load Data 

Data Source: Wisconsin Election Commission Website 
- https://elections.wi.gov/sites/default/files/documents/Statewide%20Summary%20Results_0.xlsx
- https://elections.wi.gov/sites/default/files/documents/County%20by%20County%20Report_Representative%20to%20the%20Assembly_0.xlsx
- https://elections.wi.gov/sites/default/files/documents/County%20by%20County%20Report_State%20Senator_0.xlsx

The County by County spreadsheets contain individual tabs for each contest, with
a list of results from each county that cast votes in that contest. The first
tab of information has a formatting error and needs to be fixed individually. 


```{r}
path <- "./WI_County_StateSenators.xlsx"
sen_data <- read_all_sheets(path)

path2 <- "./WI_County_Representatives_xlsx.xlsx"
rep_data <- read_all_sheets(path2)

path3 <- "./WI_Statewide_Results.xlsx"
state_data <- read_excel(path3, sheet = 2, trim_ws = T, range = "A5:L1400")

```


# Clean state senate & state representative data 

## Loop through data 

# -- State Senate races
```{r}
# Get list of state senate contests for 2022
senoffice_list <- list()

for (i in 1:length(sen_data)) {
  val <- pull(sen_data[[i]][2,][1])
  senoffice_list <- append(senoffice_list, values = val)
}

senoffice_list[[1]] <- pull(sen_data[[1]][3,][1]) ## Fix the value from the first sheet

unlist(senoffice_list) ## -- > List of all the state senate contests of that year 

```

```{r}
# Change column names
for (i in 2:length(sen_data)) {
  names(sen_data[[i]]) <- sen_data[[i]][4,] ## Get column names from the 4th row
}
# Fix first sheet 
names(sen_data[[1]]) <- sen_data[[1]][5,] ## Get column names from the 5th row
```


# -- State Representative races
```{r}
# Get list of state respresentative contests 
repoffice_list <- list()

for (i in 1:length(rep_data)) {
  val <- pull(rep_data[[i]][2,][1])
  repoffice_list <- append(repoffice_list, values = val)
}


# Fix the first NA
repoffice_list[[1]] <- pull(rep_data[[1]][3,][1])

unlist(repoffice_list) ## -- > List of all the state representative contests of that year 

```

```{r}
# Change column names for each data frame 
for (i in 2:length(rep_data)) {
  names(rep_data[[i]]) <- rep_data[[i]][4,] ## Get column names from the 4th row
}

# Fix first sheet 
names(rep_data[[1]]) <- rep_data[[1]][5,] ## Get column names from the 5th row
```


# List county-contest pairs
## -- State Senate 
```{r}
# Make dataset of all county-contest pairs for each state senate contest
sencounty_list <- list()

for (i in 1:length(sen_data)) {
  
  ## Create a list of dataframes from each sheet in the State Senate Excel
  
  df <- unique(sen_data[[i]]["County"])
  df <- df %>% filter(grepl("Total", County) == F & grepl("County", County) == F 
              & grepl("Totals:", County) == F 
              & is.na(County)==F 
              & !(County %in% senoffice_list))
  
  df["Contest_name"] <- senoffice_list[i]  

  sencounty_list[[i]] <- df
  
}
```


```{r}

## Rbind list of data frames 
sencounty_key <- do.call(rbind, sencounty_list)  ## Results in some duplicate rows 

## Make list of counties 
  WI_county_list <- sencounty_key %>% select(County) %>% filter(grepl("STATE", County)==F) %>%
    unique() %>% pull()
  
## Filter out  duplicate rows
  sencounty_key <- 
    sencounty_key %>% filter(!(County %in% senoffice_list), !(Contest_name %in% WI_county_list))

```

## -- State Representatives 

```{r}
# Make dataset of all county-contest pairs for each state representative contest
repcounty_list <- list()

for (i in 1:length(rep_data)) {
  
  ## Create a list of dataframes from each sheet in the County State Representatives Excel
  
  df <- unique(rep_data[[i]]["County"])
  df <- df %>% filter(grepl("Total", County) == F & grepl("County", County) == F 
              & grepl("Totals:", County) == F 
              & is.na(County)==F 
              & !(County %in% repoffice_list))
  
  df["Contest_name"] <- repoffice_list[i]  

  repcounty_list[[i]] <- df
  
}

```


```{r}

## Rbind list of dataframes 
repcounty_key <- do.call(rbind, repcounty_list)  ## Results in some duplicate rows 


## Filter out  duplicate rows
  repcounty_key <- 
    repcounty_key %>% 
    filter(!(County %in% repoffice_list), !(Contest_name %in% WI_county_list))

```


# List Candidate-contest pairs

## -- Senate 
```{r}

# Make dataframe of candidate - contest pairs for each contest 

sencan_list <- list()
can_list <- list()

for (i in 1:length(sen_data)) {
  
  # Loop through each contest to get a list of candidates
  # Add candidate-contest pairs to a new list of data frames 
  
  can_list <- list() ## Reset list of candidates for contest i
  
  ## Get candidate names and parties for contest i
  df <- sen_data[[i]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][5,] }
  
 
  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list)  ## Make data frame for contest i

  names(ndf)[2] <- "Candidate_name"   ## Rename candidate list column
  ndf["Contest_name"] <- senoffice_list[i]   ## Add contest column
  
  sencan_list[[i]] <- ndf ## Add contest i data frame to list of all data frames
  
}

```

```{r}
# Fill in the first contest separately due to Excel formatting error

  can_list <- list() # Reset can_list
  df <- sen_data[[1]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][6,] }

  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list) ## Make data frame for contest 1
  

  names(ndf)[2] <- "Candidate_name"   ## Rename can_list 
  ndf["Contest_name"] <- senoffice_list[1]    ## Add office column
  sencan_list[[1]] <- ndf ## Add to list of data frames

sencan_list[[1]] ## Check data frame

```

```{r}
# Rbind all state senate contests

sencan_list <- do.call(rbind, sencan_list)

```

# -- Rep 

```{r}
# Make dataframe of candidate - contest pairs for each contest 

repcan_list <- list()
can_list <- list()

for (i in 1:length(rep_data)) {
  
  # Loop through each contest to get a list of candidates
  # Add candidate-contest pairs to a new list of data frames 
 
  can_list <- list()  ## Reset list of candidates for contest i
  
  ## Get candidate names and parties for contest i
  df <- rep_data[[i]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][5,] }
  
  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list)  ## Make data frame for contest i
  
  names(ndf)[2] <- "Candidate_name" ## Rename can_list column
  ndf["Contest_name"] <- repoffice_list[i]   ## Add office column
  repcan_list[[i]] <- ndf ## Add to list of data frames
  
}

```

```{r}
# Fill in the first contest separately due to Excel formatting error

  can_list <- list() # Reset can_list
  df <- rep_data[[1]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][6,] }
  
  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list)  ## Make data frame for contest 1

  names(ndf)[2] <- "Candidate_name"   ## Rename can_list 
  ndf["Contest_name"] <- repoffice_list[1]    ## Add office column
  repcan_list[[1]] <- ndf ## Add to list of data frames

```

```{r}
# Rbind all state representative contests 

repcan_list <- do.call(rbind, repcan_list)

```


## Join county data to candidate data
```{r}
# State Senate races
senators_final <- right_join(sencounty_key, sencan_list, by = "Contest_name")

# State Representative races 
reps_final <- right_join(repcounty_key, repcan_list, by = "Contest_name")

```





