---
title: "WI_Data"
output: html_document
date: "2025-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries 

```{r}
library(tidyverse)
library(tidyr)
library(readxl)
library(stringr)
```


# Define functions 

```{r}
# Code from 
# https://www.geeksforgeeks.org/how-to-read-a-xlsx-file-with-multiple-sheets-in-r/

read_all_sheets <- function(file) {
  
  # This function reads each sheet in an inputted Excel file and drops the first
  # tab in the Excel spreadsheet.
  
  # Inputs: An Excel file with multiple sheets
  # Outputs: A list of data frames for each sheet
  
  sheets <- excel_sheets(file)  ## Get info about sheets 
  sheets <- sheets[-1] ## Get rid of "Document map" sheet
  tibble <- lapply(sheets, function(x) read_excel(file, sheet = x)) ## Read each sheet
  data_frame <- lapply(tibble, as.data.frame) ## Create data frame from each sheet
  
  
  names(data_frame) <- sheets ## Assign names to each data frame in list 
  
  return(data_frame)  ## Return list of data frame 
}

```

# Load Data 

Data Source: Wisconsin Election Commission Website 
- https://elections.wi.gov/sites/default/files/documents/Statewide%20Summary%20Results_0.xlsx
- https://elections.wi.gov/sites/default/files/documents/County%20by%20County%20Report_Representative%20to%20the%20Assembly_0.xlsx
- https://elections.wi.gov/sites/default/files/documents/County%20by%20County%20Report_State%20Senator_0.xlsx

The County by County spreadsheets contain individual tabs for each contest, with
a list of results from each county that cast votes in that contest. The first
tab of information has a formatting error and needs to be fixed individually. 


```{r}
path <- "./WI_County_StateSenators.xlsx"
sen_data <- read_all_sheets(path)

path2 <- "./WI_County_Representatives_xlsx.xlsx"
rep_data <- read_all_sheets(path2)

path3 <- "./WI_Statewide_Results.xlsx"
state_data <- read_excel(path3, sheet = 2, trim_ws = T, range = "A5:L1400")

```


# Clean state senate & state representative data 

## Loop through data 

## --- State Senate races
```{r}
# Get list of state senate contests for 2022
senoffice_list <- list()

for (i in 1:length(sen_data)) {
  val <- pull(sen_data[[i]][2,][1])
  senoffice_list <- append(senoffice_list, values = val)
}

senoffice_list[[1]] <- pull(sen_data[[1]][3,][1]) ## Fix the value from the first sheet

unlist(senoffice_list) ## -- > List of all the state senate contests of that year 

```

```{r}
# Change column names
for (i in 2:length(sen_data)) {
  names(sen_data[[i]]) <- sen_data[[i]][4,] ## Get column names from the 4th row
}
# Fix first sheet 
names(sen_data[[1]]) <- sen_data[[1]][5,] ## Get column names from the 5th row
```


## --- State Representative races
```{r}
# Get list of state respresentative contests 
repoffice_list <- list()

for (i in 1:length(rep_data)) {
  val <- pull(rep_data[[i]][2,][1])
  repoffice_list <- append(repoffice_list, values = val)
}


# Fix the first NA
repoffice_list[[1]] <- pull(rep_data[[1]][3,][1])

unlist(repoffice_list) ## -- > List of all the state representative contests of that year 

```

```{r}
# Change column names for each data frame 
for (i in 2:length(rep_data)) {
  names(rep_data[[i]]) <- rep_data[[i]][4,] ## Get column names from the 4th row
}

# Fix first sheet 
names(rep_data[[1]]) <- rep_data[[1]][5,] ## Get column names from the 5th row
```


# List county-contest pairs
## --- State Senators 
```{r}
# Make dataset of all county-contest pairs for each state senate contest
sencounty_list <- list()

for (i in 1:length(sen_data)) {
  
  ## Create a list of dataframes from each sheet in the State Senate Excel
  
  df <- unique(sen_data[[i]]["County"])
  df <- df %>% filter(grepl("Total", County) == F & grepl("County", County) == F 
              & grepl("Totals:", County) == F 
              & is.na(County)==F 
              & !(County %in% senoffice_list))
  
  df["Contest_name"] <- senoffice_list[i]  

  sencounty_list[[i]] <- df
  
}
```


```{r}

## Rbind list of data frames 
sencounty_key <- do.call(rbind, sencounty_list)  ## Results in some duplicate rows 

## Make list of counties 
  WI_county_list <- sencounty_key %>% select(County) %>% filter(grepl("STATE", County)==F) %>%
    unique() %>% pull()
  
## Filter out  duplicate rows
  sencounty_key <- 
    sencounty_key %>% filter(!(County %in% senoffice_list), !(Contest_name %in% WI_county_list))

```

## --- State Representatives 

```{r}
# Make dataset of all county-contest pairs for each state representative contest
repcounty_list <- list()

for (i in 1:length(rep_data)) {
  
  ## Create a list of dataframes from each sheet in the County State Representatives Excel
  
  df <- unique(rep_data[[i]]["County"])
  df <- df %>% filter(grepl("Total", County) == F & grepl("County", County) == F 
              & grepl("Totals:", County) == F 
              & is.na(County)==F 
              & !(County %in% repoffice_list))
  
  df["Contest_name"] <- repoffice_list[i]  

  repcounty_list[[i]] <- df
  
}

```


```{r}

## Rbind list of dataframes 
repcounty_key <- do.call(rbind, repcounty_list)  ## Results in some duplicate rows 


## Filter out  duplicate rows
  repcounty_key <- 
    repcounty_key %>% 
    filter(!(County %in% repoffice_list), !(Contest_name %in% WI_county_list))

```


# List Candidate-contest pairs

## --- State Senators
```{r}

# Make dataframe of candidate - contest pairs for each contest 

sencan_list <- list()
can_list <- list()

for (i in 1:length(sen_data)) {
  
  # Loop through each contest to get a list of candidates
  # Add candidate-contest pairs to a new list of data frames 
  
  can_list <- list() ## Reset list of candidates for contest i
  
  ## Get candidate names and parties for contest i
  df <- sen_data[[i]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][5,] }
  
 
  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list)  ## Make data frame for contest i

  names(ndf)[2] <- "Candidate_name"   ## Rename candidate list column
  ndf["Contest_name"] <- senoffice_list[i]   ## Add contest column
  
  sencan_list[[i]] <- ndf ## Add contest i data frame to list of all data frames
  
}

```

```{r}
# Fill in the first contest separately due to Excel formatting error

  can_list <- list() # Reset can_list
  df <- sen_data[[1]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][6,] }

  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list) ## Make data frame for contest 1
  

  names(ndf)[2] <- "Candidate_name"   ## Rename can_list 
  ndf["Contest_name"] <- senoffice_list[1]    ## Add office column
  sencan_list[[1]] <- ndf ## Add to list of data frames

sencan_list[[1]] ## Check data frame

```

```{r}
# Rbind all state senate contests

sencan_list <- do.call(rbind, sencan_list)

```

## --- Representatives

```{r}
# Make dataframe of candidate - contest pairs for each contest 

repcan_list <- list()
can_list <- list()

for (i in 1:length(rep_data)) {
  
  # Loop through each contest to get a list of candidates
  # Add candidate-contest pairs to a new list of data frames 
 
  can_list <- list()  ## Reset list of candidates for contest i
  
  ## Get candidate names and parties for contest i
  df <- rep_data[[i]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][5,] }
  
  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list)  ## Make data frame for contest i
  
  names(ndf)[2] <- "Candidate_name" ## Rename can_list column
  ndf["Contest_name"] <- repoffice_list[i]   ## Add office column
  repcan_list[[i]] <- ndf ## Add to list of data frames
  
}

```

```{r}
# Fill in the first contest separately due to Excel formatting error

  can_list <- list() # Reset can_list
  df <- rep_data[[1]] 
  Candidate_party <- data.frame(names(df))
  Candidate_party <- pull(filter(Candidate_party, (names.df. %in% c("DEM", "REP", "UNA", "LIB", "GRE"))))
  for (x in 1:length(Candidate_party)) { can_list[[x]] <- df[Candidate_party[x]][6,] }
  
  can_list <- unlist(can_list)
  ndf <- data.frame(Candidate_party, can_list)  ## Make data frame for contest 1

  names(ndf)[2] <- "Candidate_name"   ## Rename can_list 
  ndf["Contest_name"] <- repoffice_list[1]    ## Add office column
  repcan_list[[1]] <- ndf ## Add to list of data frames

```

```{r}
# Rbind all state representative contests 

repcan_list <- do.call(rbind, repcan_list)

```


## --- Join county data to candidate data
```{r}
# State Senate races
senators_final <- right_join(sencounty_key, sencan_list, by = "Contest_name")

# State Representative races 
reps_final <- right_join(repcounty_key, repcan_list, by = "Contest_name")

```


# Clean Statewide electoral data

## --- Clean column names 
```{r}

## change column names

names(state_data) <- c("E1", "O", "Winner", "Office", "E2", "Num_Votes", "Perc_Total_Votes", "E3", "Candidate", "E4", "Total2", "Party & Votes")


## drop empty columns
state_data <- state_data %>% select(-c("E1", "E2", "E3", "E4", "Total2", "Winner", "O"))
```

## --- Loop to fill empty rows 

The Excel data is formatted such that the office name is not repeated down the 
first column for all of its corresponding rows of data.

```{r}
# Loop to fill empty office values 

# INITIALIZE
current <- pull(state_data[1,1])

# LOOP AND FILL VALUES
for (i in 1:nrow(state_data)) {
  
  ## Loop through office names to fill down the first column of the spreadsheet
  ## with the relevant office for each row.
  
  current_office <- pull(state_data[i,1]) 
  
  if (is.na(current_office)) {state_data[i,1] <- current}
  
  if (is.na(current_office) == F) {current <- current_office}
}

```

```{r}
# Drop NA values

state_data <- drop_na(state_data)

glimpse(state_data)

```

```{r}
# Rename columns and reorganize

state_data <- state_data %>% select(-c(Num_Votes, Perc_Total_Votes)) ## Drop vote totals 

names(state_data) <- c("Contest_name", "Candidate_name", "Candidate_party") ## Rename cols

  state_data <- state_data %>% ## Change Candidate_party values 
    mutate(Candidate_party = case_when((Candidate_party != "Republican" &  Candidate_party != "Democrat" &Candidate_party != "Independent") ~ "OTH",
    Candidate_party == "Republican" ~ "REP",
    Candidate_party == "Democrat" ~ "DEM",
    Candidate_party == "Independent" ~ "IND"))
```


## --- Add counties for each congressional district  

Counties corresponding to each congressional electoral district were not 
available in the statewide dataset. Using data from Daves Redistricting, county
names were manually added based on the district map for WI congressional races
for 2022. 

Source: https://davesredistricting.org/maps#viewmap::aa64c8f5-837c-474f-819f-6eaa1094d776

```{r}


# Manually list counties for each representative district

c1 <- c("WALWORTH", "ROCK", "KENOSHA", "RACINE", "MILWAUKEE") ## District 1
c2 <- c("SAUK", "DANE", "GREEN", "LAFAYETTE", "IOWA", "ROCK") ## District 2
c3 <- c("PIERCE", "DUNN", "PEPIN", "BUFFALO", "EAU CLAIRE", "JACKSON", ## District 3
        "CHIPPEWA", "LA CROSS", "TREMPEALEAU", "MONROE", "VERNON", "CRAWFORD",
        "GRANT", "RICHLAND", "SAUK", "JUNEAU", "ADAMS", "WOOD", "PORTAGE")
c4 <- "MILWAUKEE" ## District 4
c5 <- c("DODGE", "JEFFERSON", "WAUKESHA", "WASHINGTON", "MILWAUKEE") ## District 5
c6 <- c("WAUSHARA", "WINNEBAGO", "MARQUETTE", "COLUMBIA", "GREEN LAKE", ## District 6
        "DODGE", "FOND DU LAC", "CALUMET", "MANITOWOC", "SHEBOYGAN", "OZAUKEE")
c7 <- c("ASHLAND", "BAYFIELD", "DOUGLAS", "BURNETT", "POLK", "IRON", "VILAS", ## District 7 
        "ONEIDA", "FOREST", "LANGLADE", "MARATHON", "FLORENCE", "LINCOLN", 
        "WOOD", "JUNEAU", "MONROE", "JACKSON", "CLARK", "TAYLOR", "CHIPPEWA",
        "RUSK", "BARRON", "ST. CROIX", "WASHBURN", "SAWYER", "PRICE", "TAYLOR") 
c8 <- c("WAUPACA", "WINNEBAGO", "CALUMENT", "BROWN", "OUTAGAMIE", "MENOMINEE",## District 8
        "OCONTO", "DOOR", "KEWAUNEE", "MARINETTE")


# Create master WI County list 

WI_county_list_all <- unique(c(c1, c2, c3, c4, c5, c6, c7, c8)) 

```

```{r}

# Make data frame of county-congressional district pairs

## Make list of all congressional races for 2022
repraces <- state_data %>% select(Contest_name) %>% unique() %>% 
  filter( grepl("REPRESENTATIVE IN CONGRESS", Contest_name)==T) %>% pull()


## Repeat congressional races for each county in district

rep_office_list <- list()

for (i in 1:8) {
  
  ## Loop through county lists to get the number of times each congressional 
  ## district should be repeated in dataset
  
  variable_name <- paste0("c", i)
  replength <- length(eval(parse(text = variable_name)))
  
  rep_office_list <- append(rep_office_list, rep(repraces[i], replength))
}

# Create data set of County-Congressional Race pairs
statewiderep_key <- data.frame(County = c(c1, c2, c3, c4, c5, c6, c7, c8),
                               Contest_name = unlist(rep_office_list))


```

## --- Add counties to statewide offices

```{r}
# Get list of statewide offices
stateoffice_list <- state_data %>% select(Contest_name) %>% unique() %>% 
  filter(!(Contest_name %in% senoffice_list) & !(Contest_name %in% repoffice_list) & grepl("REP", Contest_name)==F) %>% pull()

# Repeat statewide offices 72 times (once for each county)
state_office_list <- list()
for (i in 1:length(stateoffice_list)) {state_office_list <- append(state_office_list, rep(stateoffice_list[i], length(WI_county_list_all)))}

# Create data set of County-statewide office pairs
statewide_key <- data.frame(County = rep(WI_county_list_all, length(stateoffice_list)),
                           Contest_name = unlist(state_office_list))
```

## --- Join statewide data to candidate data

```{r}
#Statewide data key 
state_data_final <- left_join(rbind(statewiderep_key, statewide_key), state_data, by = "Contest_name")

```


# Join statewide, state representative, and state senator contest data

```{r}
data_final <- rbind(reps_final, senators_final, state_data_final)
data_final <- data_final %>% unique()
```

# Check & fix data for export 

## --- Check and fix county names

```{r}
# Typos found in data

## CALUMENT ~ CALUMET 
## LA CROSS ~ LA CROSSE

data_final <- data_final %>% mutate(County = ifelse(County == "LA CROSS", "LA CROSSE", County),
                                County = ifelse(County == "CALUMENT", "CALUMET", County))

```

## --- View final data 

```{r}
data_final
```


## --- Export csvs of all keys 

```{r}
# Counties and State Senate Districts 
## write_csv(sencounty_key, "../Exports/WI_SenDistricts_Key.csv")

# State Senate Districts and Candidates
## write_csv(sencan_list, "../Exports/WI_SenCandidate_Key.csv")

# Counties and State Representative Districts
## write_csv(repcounty_key, "../Exports/WI_RepDistricts_Key.csv")

# State Representative Districts and Candidates 
## write_csv(repcan_list, "../Exports/WI_RepCandidates_Key.csv")

# Counties and Congressional Representative Districts 
## write_csv(statewiderep_key, "../Exports/WI_CongressDistricts_Key.csv")
 
# All WI counties 
## write_csv(data.frame(County_name = unique(data_final$County)), "../Exports/WI_County_List.csv")

```


## --- Export to google sheets 

```{r}
# Export data to google sheets for further visualization and analysis. 

## library(googlesheets4)
## ss <- gs4_create(
##  "Wisoncin Election Data",
##  sheets = data_final)
```




